\hspace{1cm} Do syntactic and semantic similarity lead to interference effects? \newline Evidence from self-paced reading and event-related potentials using German

### 2024-08-23 paper/ResponseToReviewers.tex

\newpage

\subsection*{Comment R1.5}
\begin{quote}
``Another problem I have is more general and has to do with progress in science. The authors in my reading find the current findings more convincing than, say, Mertzen et al. study. I get that the current study is convincing (I do find it convincing) but that does not disqualify what Mertzen et al. found in any way - both results co-exist. Ideally, I would expect that a model would be able to explain the current data, but also the past data. I get it that this is not always possible but then, \textbf{it would be good if the authors make their position explicit - either by admitting that only some data can be captured, or, if the authors believe that the past data should be ignored, it should be explained why this is so.}

All in all, I think the interpretation of the data and the model have to be strengthened to meet requirements 1-3. Alternatively, the manuscript should explicitly admit issues in the proposed interpretation - that is, explain that the model fails to capture encoding interference (point 1 if I am right), explain why the model underpredicts interference in syntax/interaction and explain why it is enough, at least for now, to only model the current findings and to put previous findings aside (assuming they actually differ - see point 3 about Bayes factors above).''\end{quote}

\subsubsection*{Response to R1.5}
About why our results diverge from previous studies' and whether we think our results are correct and the previous work is not, we have no position on this. We are simply reporting what we found. Contradictory results have been reported for many phenomena: e.g., predictive parsing effects \citep{nieuwland_etal_2018, Delong_etal_2005}, reflexives processing \citep{jaeger_etal_2020, dillon2013}, main verb/reduced relative garden paths \citep{trueswell1994semantic, ferreira1986independence}, local coherence \citep{lcpaape2023, taboretal04}, locality effects in German \citep{levy&keller_2013, vasishth2018_signficancefilter}, interference from extra-sentential distractors \citep{vandyke_mcelree06, mertzen_repVDM2006}, attenuation of agreement attraction due to case marking \citep{hartsuiker2001_agree, Avetisyan_etal_2020_agree} and relative clauses in Chinese \citep{hsiao03, gibsonwu, VasishthetalPLoSOne2013, jaeger2015subject}. The only way to resolve contradictions in the literature is by carrying out adequately powered studies in the future that can distinguish between the theoretical claims. Our work does not stand out as an anomaly; if anything, it's quite the opposite. We have added a discussion of divergent results in the literature to page \pageref{not_an_anomaly} in the General Discussion:

### 2024-08-24 paper/main.tex

\textcolor{blue}{For their recent simulations, \textcite{mertzen} created a version of the Lewis and Vasishth (2005) model that assumes that in the \textcite{vandyke07} design, three cues are used at the verb: \{$\pm$ animacy\}, \{$\pm$ grammatical subject\} and \{$\pm$ same clause\}. The \{$\pm$ same clause\} cue matches only with nouns that occurred in the same clause as the verb. ``The addition of the \{$\pm$ same clause\} cue [was] necessary to identify the correct subject'' and as a means for the model ``to achieve correct retrieval'' in the condition with high syntactic and high semantic interference \citep[][p. 14]{mertzen}. This model predicted approximately equally sized effects of syntactic and semantic interference. Although it is clear that our data does not support equally sized effects of syntactic and semantic interference, we decided to include their version of the model here because it has previously been claimed to be a good model for the present design. We have implemented two alternative versions of the model. Given the evidence against syntactic interference in our data, it is plausible that, at least in this experimental design and language, syntactic interference plays little to no role. We speculate that this could be due to the parser tracking hierarchical structure and that this hierarchical information is used when searching for a noun in memory -- even more so than suggested by \textcite{mertzen}.} \Copy{clause_tracker}{\textcolor{blue}{So, we implemented a model which puts even more emphasis on hierarchical syntactic structure and thereby basically eliminates syntactic interference.} Specifically, the relevant syntactic cues that might be used at the verb are not \{$\pm$ grammatical subject\} and \{$\pm$ same clause\} -- as implemented by \textcite{mertzen} -- but rather a composite cue \{$\pm$ subject-in-same-clause\}. \textcolor{blue}{As already noted by \textcite{mertzen}, structural information like [+ subject of clause$_i$] is not an inherent feature like for example [+ animate], but within in the cue-based retrieval framework, as a simplification, we assume that such a feature is encoded into memory and can consequently be used to match the retrieval cue \{$\pm$ subject of clause$_i$\}. This idea goes back to an assumption in the Lewis and Vasishth (2005) model by making the model ``sensitive to whether or not it is parsing an embedded clause'' \citep[][p.388]{Lewis2005} and implies a simple clause tracking mechanism.} The resulting model assumes two retrieval cues: \{$\pm$ subject-in-same-clause\} and \{$\pm$ animate\}. This proposal is \textcolor{blue}{furthermore} in the spirit of the \textcite{dillon2013} and \textcite{Sturt2003} claim, that the parser intelligently targets only the syntactically relevant noun during antecedent retrieval. In the present case, it is possible that syntactically relevant noun is only the one in the same clause as the verb. \textcolor{blue}{For more discussion of the role of structural retrieval cues, see \textcite{franck2020hierarchical} and \textcite{arnett2017subject}.}}\label{clause_tracker} \textcolor{blue}{The proposed model here using two cues predicts no syntactic interference and no interaction, but does predict semantic interference effects. Lastly, we implemented a model with one cue which unambiguously matches the subject. This cue would be the master composite cue \{$\pm$ animate-subject-in-same-clause\}. This model naturally does not predict any interference effects. It can be regarded the null model or no-interference model.}

@@

\Copy{not_an_anomaly}{\textcolor{blue}{While it seems unfortunate that the present and previous results together do not create a clear and unambiguous picture of the investigated phenomena, this divergence of results is not an anomaly within the scientific literature \citep[see e.g.,][for a high-powered failed replication attempt of a classic result in the area of predictive processing]{nieuwland_etal_2018}. The only way to resolve contradictions in the literature -- like the contradictory results between the present and previous studies regarding syntactic interference -- is by carrying out adequately powered studies in the future to reach consensus over time.}}\label{not_an_anomaly}

>>>>

\textcolor{blue}{For their recent simulations, \textcite{mertzen} created a version of the Lewis and Vasishth (2005) model that assumes that in the \textcite{vandyke07} design, three cues are used at the verb: \{$\pm$ animacy\}, \{$\pm$ grammatical subject\} and \{$\pm$ same clause\}. The \{$\pm$ same clause\} cue matches only with nouns that occurred in the same clause as the verb. ``The addition of the \{$\pm$ same clause\} cue [was] necessary to identify the correct subject'' and as a means for the model ``to achieve correct retrieval'' in the condition with high syntactic and high semantic interference \citep[][p. 14]{mertzen}. This model predicted approximately equally sized effects of syntactic and semantic interference. Although it is clear that our data does not support equally sized effects of syntactic and semantic interference, we decided to include their version of the model here because it has previously been claimed to be a good model for the present design. We have implemented two alternative versions of the model. Given the evidence against syntactic interference in our data, it is plausible that, at least in this experimental design and language, syntactic interference plays little to no role. We speculate that this could be due to the parser tracking hierarchical structure and that this hierarchical information is used when searching for a noun in memory -- even more so than suggested by \textcite{mertzen}.} \Copy{clause_tracker}{\textcolor{blue}{So, we implemented a model which puts even more emphasis on hierarchical syntactic structure and thereby basically eliminates syntactic interference.} Specifically, the relevant syntactic cues that might be used at the verb are not \{$\pm$ grammatical subject\} and \{$\pm$ same clause\} -- as implemented by \textcite{mertzen} -- but rather a composite cue \{$\pm$ subject-in-same-clause\}. \textcolor{blue}{As already noted by \textcite{mertzen}, structural information like [+ subject of clause$_i$] is not an inherent feature like for example [+ animate], but within in the cue-based retrieval framework, as a simplification, we assume that such a feature is encoded into memory and can consequently be used to match the retrieval cue \{$\pm$ subject of clause$_i$\}. This idea goes back to an assumption in the Lewis and Vasishth (2005) model by making the model ``sensitive to whether or not it is parsing an embedded clause'' \citep[][p.388]{Lewis2005} and implies a simple clause tracking mechanism.} The resulting model assumes two retrieval cues: \{$\pm$ subject-in-same-clause\} and \{$\pm$ animate\}. This proposal is \textcolor{blue}{furthermore} in the spirit of the \textcite{dillon2013} and \textcite{Sturt2003} claim, that the parser intelligently targets only the syntactically relevant noun during antecedent retrieval. In the present case, it is possible that syntactically relevant noun is only the one in the same clause as the verb. \textcolor{blue}{For more discussion of the role of structural retrieval cues, see \textcite{franck2020hierarchical} and \textcite{arnett2017subject}.}}\label{clause_tracker} \textcolor{blue}{The proposed model here using two cues predicts no syntactic interference and no interaction, but does predict semantic interference effects. Lastly, we implemented a model with one cue which unambiguously matches the subject. This cue could be the composite cue \{$\pm$ animate-subject-in-same-clause\} or it could be even more specific. This model naturally does not predict any interference effects. It can be regarded the null model or no-interference model.}

@@

\Copy{not_an_anomaly}{\textcolor{blue}{While it seems unfortunate that the present and previous results together do not create a clear and unambiguous picture of the investigated phenomena, this divergence of results is not an anomaly within the scientific literature. Contradictory results have been reported for many phenomena: e.g., predictive parsing effects \citep{nieuwland_etal_2018, Delong_etal_2005}, reflexives processing \citep{jaeger_etal_2020, dillon2013}, main verb/reduced relative garden paths \citep{trueswell1994semantic, ferreira1986independence}, local coherence \citep{lcpaape2023, taboretal04}, locality effects in German \citep{levy&keller_2013, vasishth2018_signficancefilter}, interference from extra-sentential distractors \citep{vandyke_mcelree06, mertzen_repVDM2006}, attenuation of agreement attraction due to case marking \citep{hartsuiker2001_agree, Avetisyan_etal_2020_agree} and relative clauses in Chinese \citep{hsiao03, gibsonwu, VasishthetalPLoSOne2013, jaeger2015subject}. The only way to resolve contradictions in the literature -- like the contradictory results between the present and previous studies regarding syntactic interference -- is by carrying out adequately powered studies in the future to reach consensus over time.}}\label{not_an_anomaly}
